---
title: |
  |
  | GENOME SEQUENCING REPORT
  |
subtitle: "Output by the **venae** pipeline"
date: "`r Sys.Date()`"
pagetitle: "`r paste(params$runname, params$hour)`"

output:
  bookdown::html_document2:
    keep_md: no
    code_folding: hide
    theme: flatly
    toc: yes
    toc_float: yes
    toc_depth: 2
    self_contained: true

params:
  reads: "x"
  failed: "x"
  staramr: "x"
  amrfinder: "x"
  kmer: "x"
  cgekey: "x"
  fungi: "x"
---

```{r setup_amr, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(formattable)
library(kableExtra)
library(tibble)

if(file.exists(params$failed) && file.size((params$failed)) != 0){
  failed <- read_tsv((params$failed), col_names = "sample", col_types = "c")
  failed_samples <- failed %>% pull(sample)
} else {
  failed_samples <- c()
}

fungi <- import_fungi_list((params$fungi))

```

&nbsp;

## ANTIMICROBIAL RESISTANCE  {-}


`r if(length(failed_samples) == 0 ){"All samples had enough reads for assembly-based antimicrobial resistance gene detection."} else {paste("**NOTE**: The following sample(s) did not have enough reads for assembly-based antimicrobial resistance gene detection:", paste(unlist(failed %>% pull(sample)), collapse=", "))}`

### StarAMR (bacteria, assembly-based) {-}

```{r staramr2, message=FALSE, echo=FALSE, warning=FALSE, crop = TRUE, results='asis'}

if(!file.exists(params$staramr) || file.size((params$staramr)) == 0){

  cat(paste("Not enough data to perform assembly-based antimicrobial resistance gene detection in any sample."))

} else {

  df <- import_staramr((params$staramr))

  df2 <- df %>%
    select(sample, data, perc_ident, perc_coverage, cge_phenotype) %>%
    arrange(sample, data) %>%
    filter(!str_detect(data, "Oqx|oqx")) %>%
    filter(data != "sitABCD")

  # import read data to get complete list of samples but removing those that failed assembly
  samples <- import_sample_names((params$samplenames)) %>%
    filter(!(sample %in% failed_samples))

  # format table
  df3 <- left_join(samples, df2, by = "sample") %>%
    select(Sample = sample, Gene = data, `Percent identity (%)` = perc_ident, `Percent coverage (%)` = perc_coverage, `Predicted resistance phenotype` = cge_phenotype) %>%
    mutate(`Percent identity (%)` = color_code_gene_percentage(`Percent identity (%)`, 98)) %>%
    mutate(`Percent coverage (%)` = color_code_gene_percentage(`Percent coverage (%)`, 90)) %>%
    mutate(Gene = ifelse(is.na(Gene), "None", Gene)) %>%
    mutate(Gene = ifelse(Gene != "None" & !is.na(Gene), cell_spec(Gene, "html", bold = TRUE), cell_spec(Gene, "html"))) %>%
    filter(!((Sample %in% fungi) & Gene == "None"))

  # add footnotes
  names(df3)[4] <- paste0(names(df3)[4], footnote_marker_symbol(1))
  names(df3)[3] <- paste0(names(df3)[3], footnote_marker_symbol(1))
  names(df3)[5] <- paste0(names(df3)[5], footnote_marker_symbol(2))

  df3 %>%
    # mutate(`Number of reads` = color_bar("orange")(`Number of reads`)) %>%
    kable("html", escape = "F", digits = 2, align = "l", caption = "Antimicrobial resistance genes detected by StarAMR") %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(height = "500px") %>%
    footnote(symbol = c("Percent identity and percent coverage are relative to the StarAMR reference gene. Percent identity below 98 % and percent coverage below 90% are flagged in red and should be interpreted with caution.\n", "Predicted phenotype is derived from the Centre for Genomic Epidemiology ResFinder database and does not represent all resistant determinants or point mutations. These phenotypes have not been completely validated in all organisms and may not represent the true phenotype for each antimicrobial."))
}
```

### KmerResistance (bacteria, read-based) {-}

```{r kmerres, message=FALSE, echo=FALSE, warning=FALSE, crop = TRUE, results='asis'}

if(file.size((params$kmer)) == 0){

  cat(paste("Not enough data to perform read-based antimicrobial resistance gene detection in any sample."))

} else {

  df <- read_tsv((params$kmer), col_types = "cccc__ddd__d_", col_names = c("sample", "data", "score", "expected", "p_value", "perc_ident", "perc_coverage", "depth")) %>%
  filter(!str_detect(data, "\\.")) %>%
  select(sample, data, perc_ident, perc_coverage) %>%
  arrange(sample, data)

  # import read data to get complete list of samples
  samples <- import_sample_names((params$samplenames)) %>%
    arrange(sample)

  # import CGE key
  cge_key <- read_tsv((params$cgekey), col_names = c("data", "class", "phenotype", "PMID", "mechanism", "notes", "required_gene")) %>%
    select(data, class, phenotype, notes)

  # join kmerRes data to CGE key to get predicted phenotype for each gene
  df2 <- left_join(df, cge_key, by = join_by(data), relationship = "many-to-many") %>%
      filter(!str_detect(data, "Oqx|oqx"))

  # format table
  df3 <- left_join(samples, df2, by = "sample") %>%
    select(Sample = sample, Gene = data, `Percent identity (%)` = perc_ident, `Percent coverage (%)` = perc_coverage, `Predicted resistance phenotype` = phenotype) %>%
    mutate(`Percent identity (%)` = color_code_gene_percentage(`Percent identity (%)`, 98)) %>%
    mutate(`Percent coverage (%)` = color_code_gene_percentage(`Percent coverage (%)`, 90)) %>%
    mutate(Gene = ifelse(is.na(Gene), "None", Gene)) %>%
    mutate(Gene = ifelse(Gene != "None" & !is.na(Gene), cell_spec(Gene, "html", bold = TRUE), cell_spec(Gene, "html")))

  # add footnotes
  names(df3)[4] <- paste0(names(df3)[4], footnote_marker_symbol(1))
  names(df3)[3] <- paste0(names(df3)[3], footnote_marker_symbol(1))
  names(df3)[5] <- paste0(names(df3)[5], footnote_marker_symbol(2))

  df3 %>%
    # mutate(`Number of reads` = color_bar("orange")(`Number of reads`)) %>%
    kable("html", escape = "F", digits = 2, align = "l", caption = "Antimicrobial resistance genes detected by KmerResistance") %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(height = "500px") %>%
    footnote(symbol = c("Percent identity and percent coverage are relative to the Centre for Genomic Epidemiology ResFinder database. Percent identity below 98 % and percent coverage below 90 % are flagged in red and should be interpreted with caution.", "Predicted phenotype is derived from the Centre for Genomic Epidemiology ResFinder database and does not represent all resistant determinants or point mutations. These phenotypes have not been completely validated in all organisms and may not represent the true phenotype for each antimicrobial."))
}

```

### ChroQueTas (fungi, assembly-based) {-}

```{r chroquetas, message=FALSE, echo=FALSE, warning=FALSE, crop = TRUE, results='asis'}

# if no fungal organisms detected
if(length(fungi) == 0){

  cat(paste("No fungal organisms detected."))

} else {

  # import fungi samples to get complete list of samples
  samples <- tibble(sample = fungi)

  # If fungi in failed assemblies (not enough read coverage)
  if (length(failed_samples) > 0 && dim(subset(samples, sample %in% failed_samples))[1] != 0 ){

    # don't need to print failed samples as its already printed above
    # remove failed samples from list
    fungi_samples <- samples %>%
      filter(!(sample %in% failed_samples))

  } else {

  # if no failed fungi samples, all are valid
  fungi_samples <- samples

  }

  # If no AMR genes run on any samples
  if(!file.exists(params$chroquetas) || file.size((params$chroquetas)) == 0){

    # if there are only valid samples, print list with no AMR genes
    if(dim(fungi_samples)[1] != 0){

      cat(paste("Not enough data to perform AMR mutation detection (due to incomplete assembly) in the following isolate(s): ", paste(unlist(fungi_samples), collapse = ", "), "\n", sep = "\n\n"))

      }

  } else {

  # import chroquetas file
  df <- import_chroquetas(params$chroquetas)

    # join to non-failed sample list
    df1 <- left_join(fungi_samples, df, by = "sample")

    # If one isolate had chroquetas fail due to missing reference sequences:
    none <- df1 %>%
      filter(is.na(protein)) %>%
      pull(sample)

    if(length(none) != 0){
      cat(paste("Not enough data to perform AMR mutation detection (due to incomplete assembly) in the following isolate(s): ", paste(unlist(none), collapse = ", "), "\n", sep = "\n\n"))
    }

    # If one isolate has no AMR mutations detected:
    nohits <- df1 %>%
      filter(protein == "None") %>%
      pull(sample)

    if(length(nohits) != 0){
      cat(paste("No AMR mutations detected in the following isolate(s): ", paste(unlist(nohits), collapse = ", "), "\n", sep = "\n\n"))
    }

    # remove those with no hits
    df2 <- df1 %>%
      filter(protein != "None")

    # do not print table if there are no hits
    if(dim(df2)[1] > 0){

    # format table
    df3 <- df2 %>%
      mutate(resistance_confidence = case_when(
        str_detect(resistance_confidence, "^1") ~ cell_spec(resistance_confidence, "html", color = "#00BFC4", bold = TRUE),
        str_detect(resistance_confidence, "^2") ~ cell_spec(resistance_confidence, "html", color = "#00BFC4", bold = TRUE),
        str_detect(resistance_confidence, "^3") ~ cell_spec(resistance_confidence, "html", color = "#00BFC4", bold = TRUE),
        str_detect(resistance_confidence, "^4") ~ cell_spec(resistance_confidence, "html", color = "#00BFC4", bold = TRUE),
        str_detect(resistance_confidence, "^5") ~ cell_spec(resistance_confidence, "html", color = "#F8766D", bold = TRUE),
        str_detect(resistance_confidence, "^6") ~ cell_spec(resistance_confidence, "html", color = "#F8766D", bold = TRUE),
        str_detect(resistance_confidence, "^7") ~ cell_spec(resistance_confidence, "html", color = "#F8766D", bold = TRUE),
        str_detect(resistance_confidence, "^8") ~ cell_spec(resistance_confidence, "html", color = "#F8766D", bold = TRUE),
        TRUE ~ cell_spec(resistance_confidence, "html"))) %>%
      select(Sample = sample, Protein = protein, Mutation = mutation, `Resistance confidence score` = resistance_confidence, `Predicted phenotype` = phenotype) %>%
      mutate(Protein = ifelse(Protein != "None" & !is.na(Protein), cell_spec(Protein, "html", bold = TRUE), cell_spec(Protein, "html")))

    # add footnotes
    names(df3)[4] <- paste0(names(df3)[4], footnote_marker_symbol(1))
    names(df3)[5] <- paste0(names(df3)[5], footnote_marker_symbol(2))

    df3 %>%
      # mutate(`Number of reads` = color_bar("orange")(`Number of reads`)) %>%
      kable("html", escape = "F", digits = 2, align = "l", caption = "Antimicrobial resistance mutations detected by ChroQueTas") %>%
      kable_styling(bootstrap_options = c("striped", "hover")) %>%
      scroll_box(height = "500px") %>%
      #collapse_rows(columns = 1:3, valign = "middle") %>%
      footnote(symbol = c(
        "Confidence scores are derived from the ChroQueTas tool, and blue colour indicates scores with strong evidence (1-4), and red indicates scores with weak evidence (5-8). NA indicates no data. For more information on confidence scores, see [here](https://github.com/nmquijada/ChroQueTas/wiki/Confidence-score-for-antimicrobial-resistance).",
        "Predicted phenotype is derived from the FungAMR database and does not represent all resistant determinants or point mutations. These phenotypes have not been validated and may not represent the true phenotype for each antimicrobial. Predicted resistance should be confirmed by phenotypic susceptibility testing."))

    }
  }
}

```
