---
title: |
  |
  | GENOME SEQUENCING REPORT
  |
subtitle: "Output by the **venae** pipeline"
date: "`r Sys.Date()`"
pagetitle: "`r paste(params$runname, params$hour)`"

output:
  bookdown::html_document2:
    keep_md: no
    code_folding: hide
    theme: flatly
    toc: yes
    toc_float: yes
    toc_depth: 2
    self_contained: true

params:
  reads: "x"
  failed: "x"
  staramr: "x"
  amrfinder: "x"
  kmer: "x"
  cgekey: "x"
  fungi: "x"
---

```{r setup_amr, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(formattable)
library(kableExtra)
library(tibble)

if(file.exists(params$failed) && file.size((params$failed)) != 0){
  failed <- read_tsv((params$failed), col_names = "sample", col_types = "c")
  failed_samples <- failed %>% pull(sample)
} else {
  failed_samples <- c()
}

fungi <- import_fungi_list((params$fungi))

```

&nbsp;

## ANTIMICROBIAL RESISTANCE  {-}


`r if(length(failed_samples) == 0 ){"All samples had enough reads for assembly-based antmicrobial resistance gene detection with StarAMR and AMRFinder"} else {paste("**NOTE**: The following sample(s) did not have enough reads for assembly-based antimicrobial resistance gene detection with StarAMR and AMRFinder:", paste(unlist(failed %>% pull(sample)), collapse=", "))}`

### StarAMR (assembly-based) {-}

```{r staramr2, message=FALSE, echo=FALSE, warning=FALSE, crop = TRUE, results='asis'}

if(!file.exists(params$staramr) || file.size((params$staramr)) == 0){

  cat(paste("Not enough data to perform assembly-based antimicrobial resistance gene detection in any sample."))

} else {

  df <- import_staramr((params$staramr))

  df2 <- df %>%
    select(sample, data, perc_ident, perc_coverage, cge_phenotype) %>%
    arrange(sample, data) %>%
    filter(!str_detect(data, "Oqx|oqx")) %>%
    filter(data != "sitABCD")

  # import read data to get complete list of samples but removing those that failed assembly
  samples <- import_sample_names((params$samplenames)) %>%
    filter(!(sample %in% failed_samples))

  # format table
  df3 <- left_join(samples, df2, by = "sample") %>%
    select(Sample = sample, Gene = data, `Percent identity (%)` = perc_ident, `Percent coverage (%)` = perc_coverage, `Predicted resistance phenotype` = cge_phenotype) %>%
    mutate(`Percent identity (%)` = color_code_gene_percentage(`Percent identity (%)`, 98)) %>%
    mutate(`Percent coverage (%)` = color_code_gene_percentage(`Percent coverage (%)`, 90)) %>%
    mutate(Gene = ifelse(is.na(Gene), "None", Gene)) %>%
    mutate(Gene = ifelse(Gene != "None" & !is.na(Gene), cell_spec(Gene, "html", bold = TRUE), cell_spec(Gene, "html"))) %>%
    filter(!((Sample %in% fungi) & Gene == "None"))

  # add footnotes
  names(df3)[4] <- paste0(names(df3)[4], footnote_marker_symbol(1))
  names(df3)[3] <- paste0(names(df3)[3], footnote_marker_symbol(1))
  names(df3)[5] <- paste0(names(df3)[5], footnote_marker_symbol(2))

  df3 %>%
    # mutate(`Number of reads` = color_bar("orange")(`Number of reads`)) %>%
    kable("html", escape = "F", digits = 2, align = "l", caption = "Antimicrobial resistance genes detected by StarAMR") %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(height = "500px") %>%
    footnote(symbol = c("Percent identity and percent coverage are relative to the StarAMR reference gene. Percent identity below 98 % and percent coverage below 90% are flagged in red and should be interpreted with caution.\n", "Predicted phenotype is derived from the Centre for Genomic Epidemiology ResFinder database and does not represent all resistant determinants or point mutations. These phenotypes have not been completely validated in all organisms and may not represent the true phenotype for each antimicrobial."))
}
```

### KmerResistance (read-based) {-}

```{r kmerres, message=FALSE, echo=FALSE, warning=FALSE, crop = TRUE, results='asis'}

if(file.size((params$kmer)) == 0){

  cat(paste("Not enough data to perform read-based antimicrobial resistance gene detection in any sample."))

} else {

  df <- read_tsv((params$kmer), col_types = "cccc__ddd__d_", col_names = c("sample", "data", "score", "expected", "p_value", "perc_ident", "perc_coverage", "depth")) %>%
  filter(!str_detect(data, "\\.")) %>%
  select(sample, data, perc_ident, perc_coverage) %>%
  arrange(sample, data)

  # import read data to get complete list of samples
  samples <- import_sample_names((params$samplenames)) %>%
    arrange(sample)

  # import CGE key
  cge_key <- read_tsv((params$cgekey), col_names = c("data", "class", "phenotype", "PMID", "mechanism", "notes", "required_gene")) %>%
    select(data, class, phenotype, notes)

  # join kmerRes data to CGE key to get predicted phenotype for each gene
  df2 <- left_join(df, cge_key, by = join_by(data), relationship = "many-to-many") %>%
      filter(!str_detect(data, "Oqx|oqx"))

  # format table
  df3 <- left_join(samples, df2, by = "sample") %>%
    select(Sample = sample, Gene = data, `Percent identity (%)` = perc_ident, `Percent coverage (%)` = perc_coverage, `Predicted resistance phenotype` = phenotype) %>%
    mutate(`Percent identity (%)` = color_code_gene_percentage(`Percent identity (%)`, 98)) %>%
    mutate(`Percent coverage (%)` = color_code_gene_percentage(`Percent coverage (%)`, 90)) %>%
    mutate(Gene = ifelse(is.na(Gene), "None", Gene)) %>%
    mutate(Gene = ifelse(Gene != "None" & !is.na(Gene), cell_spec(Gene, "html", bold = TRUE), cell_spec(Gene, "html")))

  # add footnotes
  names(df3)[4] <- paste0(names(df3)[4], footnote_marker_symbol(1))
  names(df3)[3] <- paste0(names(df3)[3], footnote_marker_symbol(1))
  names(df3)[5] <- paste0(names(df3)[5], footnote_marker_symbol(2))

  df3 %>%
    # mutate(`Number of reads` = color_bar("orange")(`Number of reads`)) %>%
    kable("html", escape = "F", digits = 2, align = "l", caption = "Antimicrobial resistance genes detected by KmerResistance") %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(height = "500px") %>%
    footnote(symbol = c("Percent identity and percent coverage are relative to the Centre for Genomic Epidemiology ResFinder database. Percent identity below 98 % and percent coverage below 90 % are flagged in red and should be interpreted with caution.", "Predicted phenotype is derived from the Centre for Genomic Epidemiology ResFinder database and does not represent all resistant determinants or point mutations. These phenotypes have not been completely validated in all organisms and may not represent the true phenotype for each antimicrobial."))
}

```
